# Intro to Regression
The packages that I used for this week's exercises are as follows:
```{r library, message=FALSE}
library(tidyverse) # For general functions
library(haven) # For importing SPSS file into R
```

Inporting data:
```{r data-import}
lab2_raw <- read_sav("data/lab2data.sav");
```

## Data Screening {-}
Quick way to see if any variables have missing data

```{r screening}
apply(is.na(lab2_raw), 2, which);
```
And we can see that the observation 59 has missing data for Anxiety.  For the sake of our analysis we will just remove this observation (Which is what I believe we did in the Lab);

```{r}
lab2 <- lab2_raw %>%
  na.omit()

sum(is.na(lab2$anxiety)) # Just testing that we did remove the NA
```

## Conducting a Multiple Regression
I haven't yet found a way to have all the regression output as nicely formatted as it appears in SPSS.  The closest I could get was to save the regression output in a separate APA formatted document, which I achieved with using the function <code>apa.reg.table</code> located in the **apaTables** package.  The APA output is saved in a folder called "output" which you can access from <a href = https://github.com/ajmackay/PSY40001.git>gitHub</a> if you so desire.

### Descriptives
```{r descriptives}
table1::table1(~anxiety + raceab + socialid + selfest + CBT, data = lab2)
```

### Multiple Regression
The following table provides all the numbers needed to fill in the worksheet.
```{r multiple-regression}
lab2.lm <- lm(anxiety ~ socialid + selfest + raceab + CBT, data = lab2_raw); #Saving Linear Model in lab2.lm object

summary(lab2.lm)
```
#### Output rundown {-}
##### Residuals {-}
If the residuals are roughly centered around zero and have a similar spread on each side then the model probably fits the assumption of homoscedasticity.

##### Coefficients {-}
**Estimate:** The estimated effect (aka regression coefficient).  Tells us how much the DV increases/decreases per one point increase in the predictor (controlling for other predictors).

**Std. Error:** The standard error of the estimate (regression coefficient).  Indicates how much variation there is around the estimate.

**t-value:** Displays the test statistic.  By default, the test statistic will be the t-value from a two-sided t-test.  The larger the test statistic, the less likely that the results were from chance alone.

**Pr(>|t|):** P-value
